Polymethylation scores results

### 2021-11-04 Reports/Paper1.Rmd
more revisions post Kelly 1st round comments

# Figure `r main_fig` - Selection of samples from the Fragile Families and Child Wellbeing (FFCW) study into this analysis. N represents the number of individuals at each step in the selection procedure, M represents the number of samples, as individuals with repeated measures can have more than one sample.

@@

if(childdata %>% filter((childteen=='C' & k5f1l=='2 no') | (childteen=='T' & k6d40=='2 No' & k5f1l!='1 yes'))%>%nrow()!=nrow(modeldata)){stop('filter N and completecase data set N do not match')}

@@

model_twovisit=modeldata_wide[complete.cases(modeldata_wide), ]
```

@@

data_text=all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n), x=-0.25)%>%
  unnest(tidied)%>%
  filter(outcome=='Sustained smoking polymethylation score (newborns, cell-type controlled)' & term=='smkPreg_binaryYes'& predictors=='secondhand smoke exposure model')%>%mutate(outcome_f=outcome)

@@

forest_plot=all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
  unnest(tidied)%>%
  filter(term=='smkPreg_binaryYes' & predictors=='secondhand smoke exposure model')%>%
  filter(outcome%in% my_outcomes)%>%
    mutate(outcome_f=factor(outcome, levels=rev(my_outcomes)))%>%
  ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=ancestry))+
  geom_point(position = position_dodge(width = 1))+
  geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
  facet_grid(.~outcome_f, scales='free_x', labeller = labeller(outcome_f=outcome_labels))+
  geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
  geom_text(data=data_text, aes(x=x,label=n), position=position_dodge(width = 1), hjust=0.05)+
  theme_classic()+
  theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
  scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')+
  scale_y_discrete('', limits=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'), labels=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'))+scale_color_manual('', values=c('black', 'darkgrey'))

@@

# Supplemental Table `r sup_tab`: Associations between prenatal maternal smoking and DNA methylation summary measures in multivariable models among `r n_total` US children by age

test=rbind(global_models, longitudinal_global_models %>% mutate(childteen='Longitudinal'))%>%
  filter(predictors=='secondhand smoke exposure model')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(childteen=factor(case_when(childteen=='Age 9'~'Age 9 models', 
                             childteen=='Age 15'~'Age 15 models', 
                             childteen=='Longitudinal'~'Longitudinal models'),levels=c('Age 9 models', 'Age 15 models', 'Longitudinal models')), 
         tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

```{r}
test=rbind(longitudinal_global_models %>% mutate(ancestry='Unstratified'),
           longitudinal_local_models%>%dplyr::select(-data))%>%
  filter(predictors=='secondhand smoke exposure model')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

@@

# Supplemental Table `r sup_tab`: Coefficients for multivariable models of prenatal maternal smoke exposure regressed on selected DNA methylation summary variables (outcomes) among children in the Fragile Families and Child Wellbeing study

@@

@@

fig4b=global_roc%>%
  dplyr::select(covariates, methylation, childteen, test)%>% 
  filter(str_detect(methylation, 'SSnewbornCT') & covariates!='')%>%unnest(test)%>%
  rbind(long_global_roc%>% 
          mutate(childteen='Age 9 & Age 15', methylation=gsub('_Age_9.*', '', methylation))%>%
          dplyr::select(covariates, methylation, childteen, test)%>% 
          filter(str_detect(methylation, 'SSnewbornCT') 
                 & covariates!='')%>%
          unnest(test))%>%
  ggplot(aes(x=spec, y=sens, color=childteen, linetype=childteen))+geom_line(size=2)+
  facet_wrap(~methylation, labeller = labeller(methylation=c('SSnewbornCT_center'='Polymethylation score, \n newborn mean centered')))+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=4,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')+
  scale_color_grey('', labels=c('Age 9 only', 'Age 9 & 15', 'Age 15 only'))+
  scale_linetype('', labels=c('Age 9 only', 'Age 9 & 15', 'Age 15 only'))+
  xlab('Specificity')+ylab('Sensitivity')+
  theme(legend.position='bottom')+ coord_fixed()

auc_fig=cowplot::plot_grid(fig4a, fig4b, labels=c('A', 'B'), nrow=1, rel_widths=c(0.65, 0.35), align='h', axis='t')

@@

# Supplemental Table with AUC and p-value comparison to sustained smoking, newborn cell type controls polymethylation score

```{r}
roc_table=global_roc %>% dplyr::select(childteen, methylation, modeltype, roc, auc)
roc_base=roc_table %>% filter(modeltype=='Base model+No methylation')%>%mutate(roc_basemodel=roc)%>%dplyr::select(childteen, roc_basemodel)
roc_ss=roc_table %>% filter(modeltype=='Sustained smoking w/ cell type correction (newborn)')%>%mutate(roc_pms=roc)%>%dplyr::select(childteen, roc_pms)

roc_table1=roc_table%>%filter(!str_detect(modeltype, 'Base model'))%>%left_join(roc_ss)%>%mutate(roctest=map2(roc, roc_pms, ~roc.test(.x, .y)))%>%mutate(auc=round(auc, 2))

rct1=roc_table1%>%mutate(`pval comparison to sustained smoking score (newborns)`=map_dbl(roctest, ~round(.x$p.value, 4)))%>%dplyr::select(childteen, modeltype, auc, `pval comparison to sustained smoking score (newborns)`)

@@

```{r, eval=stand_alone}
rct1%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value (vs polymethylation score, newborns, cell type controlled)'))
```

# Supplemental Table with AUC and p-value comparison to base model with no methylation variables (i.e., using only child sex, income to poverty ratio, saliva/epithelial cell proportions and sample plate)

roc_table2=roc_table%>%filter(str_detect(modeltype, 'Base model'))%>%left_join(roc_base)%>%mutate(roctest=map2(roc, roc_basemodel, ~roc.test(.x, .y)))%>%mutate(auc=round(auc, 2))

rct2=roc_table2%>%mutate(`pval comparison to base model`=map_dbl(roctest, ~round(.x$p.value, 4)))%>%dplyr::select(childteen, modeltype, auc, `pval comparison to base model`)

rct2%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value (vs base w/ no methyl)'))

```{r, eval=stand_alone}
rct2%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value (vs base w/ no methyl)'))
```

# Question for group: add EWAS results to this paper

-   Arguments in favor: recent paper in Clinical Epigenetics included both
-   Arguments against: enough here for a compelling story; deadline of December 1st; submit and if asked for we have it available.

```{r, eval=include_ewas}

calcLambda<-function(df, type='lm'){
  if(type=='lme'){df=df%>%mutate(P.Value=p.value)}
  o<- -log10(sort(df$P.Value, decreasing = F))
  e<- -log10(ppoints(length(df$P.Value)))
  lambda<-median(o)/median(e)
  return(lambda)
}
```

```{r, eval=include_ewas}
child_m1=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'child1_TopHits.RDS'))
teen_m1=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'teen1_TopHits.RDS'))
```

```{r, include=include_ewas}
#lme_base=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'lme_ewas_base.RDS'))
#lme_smk = lme_base %>% filter(term=='smkPreg_binaryYes')%>%mutate(p.bh=p.adjust(p.value, 'hochberg'))
#lme_smk=add_annotations(lme_smk, type='lme')
lme_smk=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'lme_smoking.RDS'))
#mutate columns to get matching columns in lme_smk as in child & teen _m1
lme_smk = lme_smk %>% mutate(logFC=estimate, CI.L=conf.low, CI.R=conf.high, P.Value=p.value, adj.P.Val=p.bh)
```

```{r, eval=F}
library(variancePartition)
dream=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'dream_ewas.RDS'))
dream_smk=topTable(dream, coef='smkPreg_binaryYes', number=nrow(dream$p.value))
#dream_smk=add_annotations(dream_smk)
```

## QQ plots

```{r, eval=include_ewas}

png(file=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.png')))
par(mfrow=c(3, 1), mar=c(5.1, 4.1, 5.1, 2.1))
  qqman::qq(child_m1$P.Value, main='Age 9: Base model EWAS')
  mtext(paste0(expression(lambda), '=', round(calcLambda(child_m1), 3)), side=3)

  qqman::qq(teen_m1$P.Value, main='Age 15: Base model EWAS')
  mtext(paste0(expression(lambda), '=', round(calcLambda(teen_m1), 3)), side=3)

  qqman::qq(lme_smk$P.Value, main='Mixed effect models: Base model EWAS')
  mtext(paste0(expression(lambda), '=', round(calcLambda(lme_smk), 3)), side=3)
dev.off()

```{r, eval=stand_alone, eval=include_ewas}
par(mfrow=c(3, 1), mar=c(5.1, 4.1, 5.1, 2.1))
  qqman::qq(child_m1$P.Value, main='Age 9: Base model EWAS')
  mtext(paste0(expression(lambda), '=', round(calcLambda(child_m1), 3)), side=3)

  qqman::qq(teen_m1$P.Value, main='Age 15: Base model EWAS')
  mtext(paste0(expression(lambda), '=', round(calcLambda(teen_m1), 3)), side=3)

  qqman::qq(lme_smk$P.Value, main='Mixed effect models: Base model EWAS')
  mtext(paste0(expression(lambda), '=', round(calcLambda(lme_smk), 3)), side=3)

## Beta vs beta plot: Age 9 vs age 15

```{r, cache=T, eval=include_ewas}
library(tibble)
base_models=list('Age_9'=child_m1%>%rownames_to_column(var='CPG'), 'Age_15'=teen_m1%>%rownames_to_column(var='CPG'), "Mixed_effect_model"=lme_smk)%>%Map(cbind, ., ModelType=names(.))

beta_base<-lapply(base_models, function(x) x %>%dplyr::select(CPG, logFC, ModelType)%>%pivot_wider(id_cols = CPG, values_from=logFC, names_from=ModelType))%>%purrr::reduce(left_join, by='CPG')

smooth_betas<-function(df, x, y){
  ggplot(df, aes_string(x=x, y=y))+
    geom_point()+geom_smooth(method='lm')+
    stat_cor()+theme_classic()+
    scale_y_continuous(name=y)+
   scale_x_continuous(name=x)+
    coord_fixed()
}

sig_beta_m1<-child_m1%>%rownames_to_column(var='CPG')%>%filter(P.Value<10e-4)%>%pull(CPG)

beta_compare=ggarrange( 
  smooth_betas(beta_base, "Age_9", "Age_15"),
  smooth_betas(beta_base %>%filter(CPG %in% sig_beta_m1), 'Age_9', 'Age_15'), 
  labels=c('A', 'B'), align='h')

beta_compare
ggexport(beta_compare, filename = file.path(outdir, paste0('SupplementalFigure', sup_fig, '.png'))) 
```

```{r, eval=stand_alone, eval=include_ewas}
beta_compare
```

## Upset plots

```{r, eval=include_ewas}
library(ggupset)
sig_cpg_list<-map(base_models, ~.x %>%filter(P.Value<10e-4)%>%dplyr::select(CPG, ModelType))%>%
  bind_rows()%>%
  group_by(CPG)%>%summarise(Models=list(ModelType))

upset_cpgs=ggplot(sig_cpg_list, aes(x=Models))+
    geom_bar() +
    scale_x_upset()
ggexport(upset_cpgs, filename = file.path(outdir, paste0('SupplementaalFigure', sup_fig, '.pdf')))
sup_fig=sup_fig+1
```

## Volcano plot

```{r, eval=include_ewas}
volcs=ggarrange(
ggplot(child_m1%>%mutate(adjPval=ifelse(adj.P.Val<0.05, '<0.05', ifelse(adj.P.Val<0.2, '<0.2', '>0.2'))), aes(x=logFC, y=-log10(P.Value), color=adjPval))+geom_point()+scale_x_continuous('Beta (regression coefficient)')+
  scale_color_manual('Benjamini-Hochberg adjusted P value', values=c('red', 'orange', 'black'))+
  scale_y_continuous(limits=c(0, 9), breaks=c(0, 2, 4, 6, 8))+
  scale_x_continuous(limits=c(-0.06, 0.06), breaks=c(-0.05, 0, 0.05)),
ggplot(teen_m1%>%mutate(adjPval=ifelse(adj.P.Val<0.05, '<0.05', ifelse(adj.P.Val<0.2, '<0.2', '>0.2'))), aes(x=logFC, y=-log10(P.Value), color=adjPval))+geom_point()+scale_x_continuous('Beta (regression coefficient)')+
  scale_color_manual('Benjamini-Hochberg adjusted P value', values=c('red', 'orange', 'black'))+
  scale_y_continuous(limits=c(0, 9), breaks=c(0, 2, 4, 6, 8))+
  scale_x_continuous(limits=c(-0.06, 0.06), breaks=c(-0.05, 0, 0.05)),
ggplot(lme_smk%>%mutate(adjPval=ifelse(adj.P.Val<0.05, '<0.05', ifelse(adj.P.Val<0.2, '<0.2', '>0.2'))), aes(x=logFC, y=-log10(P.Value), color=adjPval))+geom_point()+scale_x_continuous('Beta (regression coefficient)')+
  scale_color_manual('Benjamini-Hochberg adjusted P value', values=c('red', 'orange', 'black'))+
  scale_y_continuous(limits=c(0, 9), breaks=c(0, 2, 4, 6, 8))+
  scale_x_continuous(limits=c(-0.06, 0.06), breaks=c(-0.05, 0, 0.05)),
labels=c('A', 'B', 'C'), 
common.legend = TRUE, 
nrow=1
)

ggexport(volcs, filename=file.path(outdir, paste0('Figure', main_fig, '.pdf')))
ggexport(volcs, filename=file.path(outdir, paste0('Figure', main_fig, '.png')))
```

```{r, eval=stand_alone, eval=include_ewas, fig.width=14, }
volcs
```

## Gap hunter

```{r, eval=include_ewas}
all_gaps=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'gaphunter_gaps.RDS'))
```

## Compare across models and studies

```{r, eval=include_ewas}
topCPGs<-read.csv(paste0(datadir, "/TopCpGsWiklundEtAl2019.csv"))
topCPGs<-topCPGs %>% mutate(estimate=as.numeric(gsub("−\\s", "-", gsub(" .*", "", β..SE.))), std.error=as.numeric(sub("\\).*", "", sub(".*\\(", "", β..SE.))), p.value=P.value, source="Wiklund et al. 2019", conf.low=(estimate-1.96*std.error), conf.high=(estimate+1.96*std.error), set ='')%>%dplyr::select(CpG, Chr, Position, Gene, estimate, std.error, p.value, conf.low, conf.high, source, set)

#myCPGs<-AllModelsdf %>%filter(Model=='model1')%>%mutate(data=map(data, function(x) x %>%filter(CpG %in% sigCpGs)))%>%unnest(cols=c(data))%>%mutate(Joubertetal=ifelse(CpG %in% cpgs$CpG, 'in Joubert et al. 2016', 'Not in Joubert et al. 2016'), Wicklundetal=ifelse(CpG %in% topCPGs$CpG, 'in Wicklund et al. 2019', 'Not in Wicklund et al. 2019'))%>%mutate(Gene=gene, estimate=logFC, conf.low=CI.L, conf.high=CI.R, source=paste0('Fragile Families:', Age), std.error='NA', p.value=P.Value, Chr=chr, set='Model1', Position=pos)

```{r, fig.width=14, eval=include_ewas}
comparison=left_join(
  lme_smk%>%dplyr::select(CPG, gene, estimate, p.value)%>%dplyr::rename_with(~paste0('FF_', .x), .cols=c(estimate, p.value)), 
  cpgs%>%filter(set=='SSnewbornCT')%>%mutate(CPG=CpG)%>%dplyr::select(CPG, Gene, estimate, p.value)%>%dplyr::rename_with( ~paste0('Joubert_', .x), .cols=c(estimate, p.value))
  )%>%filter(!is.na(Joubert_estimate))

comp_nofilter=ggplot(comparison, aes(x=Joubert_estimate, y=FF_estimate, color=FF_p.value))+
  geom_point()+stat_cor()+
  scale_y_continuous('Estimate in Fragile Families')+
  scale_x_continuous('Estimate in Joubert et al. metaanalysis')+
  scale_color_continuous('P value in Fragile Families')

#comparison %>% mutate(s1=sign(FF_estimate), s2=sign(Joubert_estimate), agreement=case_when(s1==s2~'Direction consistent', s1<s2~'FF -, Joubert +', s1>s2~'FF+, Joubert-'))%>%group_by(agreement)%>%summarise(n=n())

comp_filter=ggplot(comparison%>%filter(FF_p.value<10e-4), aes(x=Joubert_estimate, y=FF_estimate, color=FF_p.value, label=ifelse(abs(Joubert_estimate)>0.04 , gene, NA)))+
  geom_point()+stat_cor()+
  geom_text_repel()+
  scale_y_continuous('Estimate in Fragile Families')+
  scale_x_continuous('Estimate in Joubert et al. metaanalysis')+
  scale_color_continuous('P value in Fragile Families')

comp_joubert=ggarrange(comp_nofilter, comp_filter, common.legend = F, labels=c('A', 'B'))
ggexport(comp_joubert, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.png')))
```

```{r, eval=stand_alone, eval=include_ewas, fig.width=14}
comp_joubert
```

## Additional covariate control and ancestry stratified models at top hits

```{r, eval=include_ewas}
top_cpgs = map(base_models, ~.x %>%filter(adj.P.Val<0.05)%>%dplyr::select(CPG, ModelType, gene))%>%bind_rows()%>%mutate(label=paste0(gene, ': ', CPG))
top_cpg_gaps=top_cpgs[top_cpgs$CPG %in% rownames(all_gaps$proberesults),]
top_cpgs=top_cpgs[!(top_cpgs$CPG %in% rownames(all_gaps$proberesults)),]
top_cpg_groups=top_cpgs %>%group_by(CPG)%>%dplyr::summarise(models=list(ModelType))
top_cpg_names=c(top_cpgs%>%pull(label)%>%unique(), 'AHRR: cg05575921', 'MYO1G: cg22132788')
top_cpgs=c(top_cpgs %>% pull(CPG)%>%unique(), 'cg05575921', 'cg22132788')

betaqc<-readRDS(file=file.path(datadir, 'OGData', "betaqc.rds"))
my_beta=as.data.frame(t(betaqc[row.names(betaqc)%in% top_cpgs, ])) %>% mutate(MethID=row.names(.))
modeldata=left_join(modeldata, my_beta)
```

```{r, eval=F}
y_vector<-top_cpgs
outcome_labels=top_cpg_names
names(outcome_labels)=y_vector

#model variables (excepting PCs, add in individually for global & ancestry specific models)
base_model_vars<-"~smkPreg_binary+cm1bsex+cm1inpov+Leukocytes_saliva+Epithelial.cells_saliva+Sample_Plate"
prenatal_model_vars<-paste0(base_model_vars, "+m1g2_YesNoPreg+m1g3_YesNoPreg")
secondhand_model_vars<-paste0(prenatal_model_vars, '+PostnatalMaternalSmokingAny+SmkAtVisitPastmonth')
interaction_model_vars<-paste0(secondhand_model_vars, '+ChildAgeComposite:smkPreg_binary')
#pc variables
global_pcs='+global_PC1+global_PC2'
local_pcs='+local_PC1+local_PC2'

#surrogate variable models
surrogate_model_vars<-paste0('~smkPreg_binary+', 
                             paste(colnames(modeldata)[str_detect(colnames(modeldata), 'sv')], collapse='+'))

predictors=c('base model'=base_model_vars, 
             'prenatal exposures model'=prenatal_model_vars, 
             'secondhand smoke exposure model'=secondhand_model_vars)

predictors_interaction=c(predictors, 'interaction model'=interaction_model_vars)

model_labels=c('base model', 'prenatal exposures model', 'secondhand smoke exposure model', 'surrogate variable models')

model_labels_interaction=c('base model', 'prenatal exposures model', 'secondhand smoke exposure model', 'interaction model (age*smoke)', 'surrogate variable models')

#stratification variables 
##############################
age_labels=c('Age 9', 'Age 15')
age_vector=modeldata$childteen%>%unique()
names(age_labels)=age_vector
ancestry_vector=modeldata$ancestry%>%unique()

#################################Source functions#######################################
model<-function(df, y, x){
  lm(formula(paste(y, x)), data=df)
}

model_lmm<-function(df, y, x){
  xnew=paste0(x, '+', '(1|id)')
  lmer(formula(paste(y, xnew)), data=df, REML=F)
}

my_clean_argnames=function(df, my_preds, childteen=T, model_labs=model_labels){
  names(model_labs)=my_preds
  df=df %>% mutate(predictors=recode_factor(predictors, !!!model_labs),
                outcome=recode_factor(outcome, !!!outcome_labels))
  if(childteen==T){df=df%>%mutate(age=recode_factor(childteen, !!!age_labels))} 
  return(df)
}

#################################Cross sectional models###########################
#########################Global
#set predictor variables & arguments
global_predictors=c(paste0(predictors, global_pcs), surrogate_model_vars)
args=list('childteen'=age_vector, 'outcome'=top_cpgs, 'predictors'=global_predictors)%>%cross_df()

#run models
global_models=modeldata%>%group_by(childteen)%>%nest()%>%left_join(args)%>%
  mutate(model=pmap(list(data, outcome, predictors), model), 
         tidied=map(model, tidy, conf.int=T), 
         glanced=map(model, glance))

#clean argument names for plotting
global_models=my_clean_argnames(global_models, global_predictors)

#########################Local
#arguments
local_predictors=c(paste0(predictors, local_pcs), surrogate_model_vars)
args=list('childteen'=c('Age 9', 'Age 15'), 'outcome'=top_cpgs, 'predictors'=local_predictors, 'ancestry'=ancestry_vector)%>%cross_df()

>>>>

# Figure `r main_fig` - Selection of samples from the Fragile Families and Child Wellbeing (FFCW) study into analytic subset. N represents the number of individuals at each step in the selection procedure, M represents the number of samples, as individuals with repeated measures can have more than one sample.

@@

if(childdata %>% filter(((k5f1l=="2 no" | (is.na(k5f1l) & k6d40=='2 No')) & childteen=='C') | (childteen=='T' & k6d40=="2 No" & (k5f1l=='2 no' |  is.na(k5f1l))))%>%nrow()!=nrow(modeldata)){stop('filter N and completecase data set N do not match')}

@@

model_twovisit=modeldata_wide[complete.cases(modeldata_wide), ]

modeldata_people=modeldata%>%
  dplyr::select(any_of(c('idnum', 'ancestry', 'cm1bsex', 'ancestry', 
                         'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'cm1inpov',
                         'smkPreg_binary', 'PostnatalMaternalSmokingAny')))%>%unique()
```

@@

data_text=all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n), x=-0.25)%>%
  unnest(tidied)%>%
  filter(outcome=='Sustained smoking polymethylation score (newborns, cell-type controlled)' & term=='smkPreg_binaryYes'& predictors=='base model')%>%mutate(outcome_f=outcome)

@@

forest_plot=all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
  unnest(tidied)%>%
  filter(term=='smkPreg_binaryYes' & predictors=='base model')%>%
  filter(outcome%in% my_outcomes)%>%
    mutate(outcome_f=factor(outcome, levels=rev(my_outcomes)))%>%
  ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=ancestry))+
  geom_point(position = position_dodge(width = 1))+
  geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
  facet_grid(.~outcome_f, scales='free_x', labeller = labeller(outcome_f=outcome_labels))+
  geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
  geom_text(data=data_text, aes(x=x,label=n), position=position_dodge(width = 1), hjust=0.05)+
  theme_classic()+
  theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
  scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')+
  scale_y_discrete('', limits=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'), labels=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'))+scale_color_manual('', values=c('black', 'darkgrey'))

@@

@@

# Supplemental Table `r sup_tab`: Coefficients for multivariable models of prenatal maternal smoke exposure regressed on selected DNA methylation summary variables (outcomes) among children in the Fragile Families and Child Wellbeing study using surrogate variables

@@

# Supplemental Table `r sup_tab`: Associations between prenatal maternal smoking and DNA methylation summary measures in multivariable linear models stratified by age or in multivariable linear mixed effect models among `r n_total` US children

test=rbind(global_models, longitudinal_global_models %>% mutate(childteen='Longitudinal'))%>%
  filter(predictors=='base model')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(childteen=factor(case_when(childteen=='Age 9'~'Age 9 models', 
                             childteen=='Age 15'~'Age 15 models', 
                             childteen=='Longitudinal'~'Longitudinal models'),levels=c('Age 9 models', 'Age 15 models', 'Longitudinal models')), 
         tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

```{r}
test=rbind(longitudinal_global_models %>% mutate(ancestry='Unstratified'),
           longitudinal_local_models%>%dplyr::select(-data))%>%
  filter(predictors=='base model')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

# Supplemental Figure `r sup_fig`: Associations between prenatal maternal smoking and polymethylation scores calculated using alternative coefficient sets among `n n_total` children in the Fragile Families and Child Wellbeing study

```{r} 
 compare_pms=all_models%>%
    mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
    unnest(tidied)%>%
    filter(term=='smkPreg_binaryYes')%>%
    filter(str_detect(outcome, 'smoking'))%>%
  filter(ancestry=='Unstratified')%>%
 ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=outcome))+
    geom_point(position = position_dodge(width = 1))+
    geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
    facet_grid(predictors~ancestry, scales='free_x')+
    geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
    theme_classic()

ggexport(compare_pms, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=9)
sup_fig=sup_fig+1

```{r, eval=stand_alone, fig.width=10}
compare_pms
```

# Supplemental Figure `r sup_fig` - Associations between prenatal maternal smoking and other *a priori* CpG sites selected from previous meta-analyses among `n n_total` children in the Fragile Families and Child Wellbeing study

```{r}
topCPGs<-read.csv(paste0(datadir, "/TopCpGsWiklundEtAl2019.csv"))
topCPGs<-topCPGs %>% mutate(estimate=as.numeric(gsub("−\\s", "-", gsub(" .*", "", β..SE.))), std.error=as.numeric(sub("\\).*", "", sub(".*\\(", "", β..SE.))), p.value=P.value, source="Wiklund et al. 2019", conf.low=(estimate-1.96*std.error), conf.high=(estimate+1.96*std.error), set ='')%>%dplyr::select(CpG, Chr, Position, Gene, estimate, std.error, p.value, conf.low, conf.high, source, set)

myCpGs=all_models%>%filter(str_detect(outcome, 'cg') & predictors=='base model')%>%unnest(tidied)%>%filter(term=='smkPreg_binaryYes')%>%
    filter(ancestry=='Unstratified')%>%mutate(CpG=gsub( '.*: ', '', outcome), Gene=gsub(':.*', '', outcome), source=paste0('Fragile Families:', age))

top_cpgs=myCpGs$CpG

cpg_all<-rbind(myCpGs%>%filter(CpG %in% top_cpgs), cpgs%>%filter(CpG %in% top_cpgs)%>%dplyr::select(-Chr), topCPGs%>%filter(CpG%in%top_cpgs)%>%mutate_at(c('p.value'), as.numeric))

plot_aprioriCpG<-ggplot(cpg_all, aes(x=estimate, xmin=conf.low, xmax=conf.high, color=source, y=CpG))+geom_point(position=position_dodge(width=1))+geom_errorbarh(height=0.1, position=position_dodge(width=1))+scale_x_continuous()+geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+facet_wrap(.~Gene, scales='free', ncol=2)+theme_classic()+theme(legend.position = 'bottom')+guides(color=guide_legend(nrow=4,byrow=TRUE))+geom_blank(data = blank_data, aes(x = estimate))+scale_color_discrete('', labels=c('Fragile families saliva: age 15', 'Fragile families saliva: age 9', 'Fragile families: mixed effect model', 'Joubert et al. 2016 newborn cord blood', 'Joubert et al. 2016 peripheral blood older children', 'Wicklund et al. peripheral blood >16 years'))

ggexport(plot_aprioriCpG, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=9)
sup_fig=sup_fig+1
```
```{r, eval=stand_alone, fig.width=10}
plot_aprioriCpG
```

@@

fig4b=global_roc%>%
  dplyr::select(covariates, methylation, childteen, test)%>% 
  filter(str_detect(methylation, 'SSnewbornCT') & covariates!='')%>%unnest(test)%>%
  rbind(long_global_roc%>% 
          mutate(childteen='Age 9 & Age 15', methylation=gsub('_Age_9.*', '', methylation))%>%
          dplyr::select(covariates, methylation, childteen, test)%>% 
          filter(str_detect(methylation, 'SSnewbornCT') 
                 & covariates!='')%>%
          unnest(test))%>%
  ggplot(aes(x=spec, y=sens, color=childteen, linetype=childteen))+geom_line(size=2)+
  facet_wrap(~methylation, labeller = labeller(methylation=c('SSnewbornCT_center_scale'='Polymethylation score, \n newborn mean centered')))+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=4,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')+
  scale_color_grey('', labels=c('Age 9 only', 'Age 9 & 15', 'Age 15 only'))+
  scale_linetype('', labels=c('Age 9 only', 'Age 9 & 15', 'Age 15 only'))+
  xlab('Specificity')+ylab('Sensitivity')+
  theme(legend.position='bottom')+ coord_fixed()

auc_fig=cowplot::plot_grid(fig4a, fig4b, labels=c('A', 'B'), nrow=1, rel_widths=c(0.65, 0.35), align='h', axis='tb')

@@

# Supplemental Table: Area under the curve for models shown in main figure 4 for classifying prenatal maternal smoke exposure, with *P* value when comparing to base model only or polymethylation score

```{r}
roc_table=global_roc %>% dplyr::select(childteen, methylation, modeltype, roc, auc)
roc_base=roc_table %>% filter(modeltype=='Base model+No methylation')%>%mutate(roc_basemodel=roc)%>%dplyr::select(childteen, roc_basemodel)
roc_ss=roc_table %>% filter(modeltype=='Sustained smoking polymethylation score (newborns, cell-type controlled)')%>%mutate(roc_pms=roc)%>%dplyr::select(childteen, roc_pms)

rct1=roc_table%>% 
  filter(methylation %in% c('globalmethylation', 'cg05575921', 'SSnewbornCT_center_scale', ''))%>%
  left_join(roc_base)%>%
  mutate(roctest_base=map2(roc, roc_basemodel, ~roc.test(.x, .y)))%>%
  left_join(roc_ss)%>%
  mutate(roctest_ss=map2(roc, roc_pms, ~roc.test(.x, .y)))%>%
  mutate(auc=round(auc, 2))%>%
  mutate(`P value vs polymethylation score`=map_dbl(roctest_ss, ~round(.x$p.value, 4)))%>%
  mutate(`P value vs base model`=map_dbl(roctest_base, ~round(.x$p.value, 4)))%>%
  dplyr::select(childteen, modeltype, auc, `P value vs base model`, `P value vs polymethylation score`)

@@

```{r, eval=stand_alone}
rct1%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value vs base model', 'P value vs polymethylation score'))
```

# Supplemental Table: Area under the curve when using different coefficients to construct polymethylation scores for predicting prenatal maternal smoke exposure

```{r} 
rct2=roc_table%>% 
    filter(str_detect(modeltype, 'smoking|No methylation'))%>%
    left_join(roc_base)%>%
    mutate(roctest_base=map2(roc, roc_basemodel, ~roc.test(.x, .y)))%>%
    left_join(roc_ss)%>%
    mutate(roctest_ss=map2(roc, roc_pms, ~roc.test(.x, .y)))%>%
    mutate(auc=round(auc, 4))%>%
    mutate(`P value vs polymethylation score with cell-type control`=map_dbl(roctest_ss, ~round(.x$p.value, 4)))%>%
    mutate(`P value vs base model`=map_dbl(roctest_base, ~round(.x$p.value, 4)))%>%
    dplyr::select(childteen, modeltype, auc, `P value vs base model`, `P value vs polymethylation score with cell-type control`)
```

```{r, eval=stand_alone}
rct2%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value vs base model', 'P value vs polymethylation score with cell-type control'))
```
